Non-word repetition in Yélî Dnye

### 2020-09-16 manuscript.Rmd

- look through this manuscript for bits of text to be written by **mc**
- do you think Figure 2 should be done with phonetic symbols? If yes, could you try to implement that?
- double check demo -- probably start with the xlsx version that has the children's first names, date of birth, and date of test. I looked again at my local files (still haven't checked the labs' ones) and I'm surprised I have maternal education for more children than I have birth order for... If you correct this, to generate the csv I also round the age, assign age.rounded to reported age (in years) when DOB is missing, remove first names, DOB, DOTest and age.precise before pushing to github
- double check my frequency entries in segments.xlsx using the search function in https://phoible.org/parameters; take a look at http://phoible.github.io/conventions/ in case you see something about double articulation (I didn't find tp or lBj)
- double check phonological implementation of stimuli in stimuli.xlsx 
- for both of the above, if you make any changes, they need to be re-saved as UTF16 txt file, and then converted (see remember above)
- double check also stim_seg_freq -- why are some items still with NA freq?
- draft discussion

@@

#quick fix, but should be fixed upstream @ac
# M2A: Seems like you already did this, and the line below
# is setting the is.na(trnsc$Sex) cases to "G" -- I assume
# that's not a desirable outcome so am commenting it out
# a2m fixed, but I think it is better to fix all of these error upstream, in the demo.xlsx file
trnsc$Sex[trnsc$Sex=="F" & !is.na(trnsc$Sex)]<-"G"
trnsc$Sex=factor(trnsc$Sex)
trnsc$birthOrder=as.numeric(trnsc$birthOrder)
```

```{r Fig1-first_vs_others, ig.pos = "T!", fig.width=5, fig.height=5, fig.cap="NWR scores for individual participants averaging separately their first attempts and all other attempts."}

@@

```{r overall}

aggregate(trnsc$correct*100,by=list(trnsc$id,trnsc$Sex,trnsc$age.rounded,trnsc$birthOrder,trnsc$matEd),mean)->overall
aggregate(trnsc$phon_score*100,by=list(trnsc$id,trnsc$Sex,trnsc$age.rounded,trnsc$birthOrder,trnsc$matEd),mean)->phon_score
aggregate(trnsc$nld*100,by=list(trnsc$id,trnsc$Sex,trnsc$age.rounded,trnsc$birthOrder,trnsc$matEd),mean)->nld
overall=cbind(overall,phon_score$x,nld$x)
colnames(overall)<-c("ID","Sex","Age","BrthO","matEd","Word","Phoneme","NLD")

@@

```{r Fig2-xling_freq, ig.pos = "T!", fig.width=5, fig.height=5, fig.cap="NWR scores for individual target words as a function of the average frequency with which each phone is found across languages."}
aggregate(trnsc$correct*100,by=list(trnsc$target,trnsc$avg_fr),mean)->xling
colnames(xling)<-c("Target","Avg_fr","Word")
plot(xling$Word~xling$Avg_fr,lty=3,ylab="NWR scores (% words)",xlab="Average phone cross-ling frequency",type="n")
text(xling$Word~xling$Avg_fr,labels=xling$Target,cex=.5)
abline(lm(xling$Word~xling$Avg_fr))
```

@@

We next checked whether this may actually be due to frequency of the sounds within the language. We estimated frequency of the phones present in the stimuli in a corpus of child-centered recordings [@casillas2020png], and applied the natural logarithm. As before, unattested sounds were not considered (i.e., they were declared NA so that they do not count for analyses). Phone frequency estimated from this corpus were correlated with cross-linguistic phone frequencies (r(`r freq_cor$parameter`)=`r round(freq_cor$estimate,3)`, =`r round(freq_cor$p.value,3)`). Therefore, we fit another mixed logistic regression, this time declaring as fixed effects scaled cross-linguistic and corpus frequencies (averaged across all attested phones within each stimulus item), in addition to age. As before, the model contained random slopes for both child ID and target. In this model, both cross-linguistic frequency (ß = `r both_coeff["avg_fr_s","Estimate"]`, SE ß = `r both_coeff["avg_fr_s","Std. Error"]`, p = `r both_coeff["avg_fr_s","Pr(>|z|)"]`) and age (ß = `r both_coeff["age.rounded_s","Estimate"]`, SE ß = `r both_coeff["age.rounded_s","Std. Error"]`, p = `r both_coeff["age.rounded_s","Pr(>|z|)"]`) were significant predictors of word-level NWR scores, but corpus frequency (ß = `r both_coeff["avg_fr_cor_ty_s","Estimate"]`, SE ß = `r both_coeff["avg_fr_cor_ty_s","Std. Error"]`, p = `r both_coeff["avg_fr_cor_ty_s","Pr(>|z|)"]`) was not.

@@

```{r bylength}
aggregate(trnsc$correct*100,by=list(trnsc$id,trnsc$age.rounded,trnsc$type),mean)->by_length
aggregate(trnsc$phon_score*100,by=list(trnsc$id,trnsc$age.rounded,trnsc$type),mean)->phon_score
aggregate(trnsc$nld*100,by=list(trnsc$id,trnsc$age.rounded,trnsc$type),mean)->nld
by_length=cbind(by_length,phon_score$x,nld$x)
colnames(by_length)<-c("ID","Age","Length","Word","Phoneme","NLD")

@@

```{r tablength}
birthorder.tab=NULL
for(myvar in c("Word","Phoneme","NLD")){
  birthorder.tab=cbind(birthorder.tab,
              paste0(round(aggregate(by_length[,myvar],by=list(by_length$Length),mean)$x),
              " (",round(aggregate(by_length[,myvar],by=list(by_length$Length),sd)$x),")"))
}
rownames(birthorder.tab)<-levels(by_length$Length)
colnames(birthorder.tab)<-c("Word","Phoneme","NLD")
apa_table(birthorder.tab,caption="NWR measured in whole-word scores, phoneme-based scores, and normalized Levenshtein Distance, separately for the four stimuli lengths.")
```

>>>>

- look through this manuscript for bits of text to be written by **mc**
- double check demo -- probably start with the xlsx version that has the children's first names, date of birth, and date of test. I looked again at my local files (still haven't checked the labs' ones) and I'm surprised I have maternal education for more children than I have birth order for... If you correct this, to generate the csv I also round the age, assign age.rounded to reported age (in years) when DOB is missing, remove first names, DOB, DOTest and age.precise before pushing to github
- double check my frequency entries in segments.xlsx using the search function in https://phoible.org/parameters; take a look at http://phoible.github.io/conventions/ in case you see something about double articulation (I didn't find tp or lBj)
- double check also stim_seg_freq -- why are some items still with NA freq?
- draft discussion

@@

#quick fix, but should be fixed upstream @ac
# M2A: Seems like you already did this, and the line below
# is setting the is.na(trnsc$Sex) cases to "G" -- I assume
# that's not a desirable outcome so am commenting it out
# a2m fixed, but I think it is better to fix all of these error upstream, in the demo.xlsx file
# M2A: Since it's fixed upstream I propose deleting it (and have therefore
# commented it out again for now)
# trnsc$Sex[trnsc$Sex=="F" & !is.na(trnsc$Sex)]<-"G"
trnsc$Sex=factor(trnsc$Sex)
trnsc$birthOrder=as.numeric(trnsc$birthOrder)
```

```{r Fig1-first_vs_others, fig.pos = "!t", fig.width=5, fig.height=5, fig.cap="NWR scores for individual participants averaging separately their first attempts and all other attempts."}

@@

```{r overall}
# M2A: This excludes a bunch of kids (anyone with NA on any of these participant features),
# whittling the summary down to 24 from 42 -- is that intended?
# Also all phon_score is 100 and all NLD is 0; does that stike you as fishy?
aggregate(trnsc$correct*100,by=list(trnsc$id,trnsc$Sex,trnsc$age.rounded,trnsc$birthOrder,trnsc$matEd),mean)->overall
aggregate(trnsc$phon_score*100,by=list(trnsc$id,trnsc$Sex,trnsc$age.rounded,trnsc$birthOrder,trnsc$matEd),mean)->phon_score
aggregate(trnsc$nld*100,by=list(trnsc$id,trnsc$Sex,trnsc$age.rounded,trnsc$birthOrder,trnsc$matEd),mean)->nld
overall=cbind(overall,phon_score$x,nld$x)
colnames(overall)<-c("ID","Sex","Age","BrthO","matEd","Word","Phoneme","NLD")

@@

```{r Fig2-xling-freq, fig.pos = "!t", fig.align = "center", out.width = '65%', fig.cap="NWR scores for individual target words as a function of the average frequency with which each phone is found across languages."}
aggregate(trnsc$correct*100,by=list(trnsc$target,trnsc$avg_fr),mean)->xling
colnames(xling)<-c("Target","Avg_fr","Word")
stims.ipa <- read_delim("stimuli2.txt", delim = "\t") %>%
  select(target, target_ortho, phono)
xling <- merge(xling, stims.ipa, by.x = "Target", by.y = "target")

item.plot <- ggplot(data = xling, aes(x = Avg_fr, y = Word, label = phono)) +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  geom_text(family = "Doulos SIL") +
  ylab("NWR scores (% words)") +
  xlab("Average phone cross-ling frequency") +
  theme_apa()
ggsave("nwr.by.freq.ITEM.png", dpi = 300, device = png())

knitr::include_graphics("nwr.by.freq.ITEM.png")

@@

We next checked whether this may actually be due to frequency of the sounds within the language. We estimated frequency of the phones present in the stimuli in a corpus of child-centered recordings [@casillas2020early], and applied the natural logarithm. As before, unattested sounds were not considered (i.e., they were declared NA so that they do not count for analyses). Phone frequency estimated from this corpus were correlated with cross-linguistic phone frequencies (r(`r freq_cor$parameter`)=`r round(freq_cor$estimate,3)`, =`r round(freq_cor$p.value,3)`). Therefore, we fit another mixed logistic regression, this time declaring as fixed effects scaled cross-linguistic and corpus frequencies (averaged across all attested phones within each stimulus item), in addition to age. As before, the model contained random slopes for both child ID and target. In this model, both cross-linguistic frequency (ß = `r both_coeff["avg_fr_s","Estimate"]`, SE ß = `r both_coeff["avg_fr_s","Std. Error"]`, p = `r both_coeff["avg_fr_s","Pr(>|z|)"]`) and age (ß = `r both_coeff["age.rounded_s","Estimate"]`, SE ß = `r both_coeff["age.rounded_s","Std. Error"]`, p = `r both_coeff["age.rounded_s","Pr(>|z|)"]`) were significant predictors of word-level NWR scores, but corpus frequency (ß = `r both_coeff["avg_fr_cor_ty_s","Estimate"]`, SE ß = `r both_coeff["avg_fr_cor_ty_s","Std. Error"]`, p = `r both_coeff["avg_fr_cor_ty_s","Pr(>|z|)"]`) was not.

@@

```{r bylength}
# M2A: this reduces the total number of kids represented from 42 to 40
aggregate(trnsc$correct*100,by=list(trnsc$id,trnsc$age.rounded,trnsc$type),mean)->by_length
aggregate(trnsc$phon_score*100,by=list(trnsc$id,trnsc$age.rounded,trnsc$type),mean)->phon_score
aggregate(trnsc$nld*100,by=list(trnsc$id,trnsc$age.rounded,trnsc$type),mean)->nld
by_length=cbind(by_length,phon_score$x,nld$x)
colnames(by_length)<-c("ID","Age","Length","Word","Phoneme","NLD")

@@

```{r tablength, results = "asis"}
birthorder.tab=NULL
for(myvar in c("Word","Phoneme","NLD")){
  birthorder.tab=cbind(birthorder.tab,
              paste0(round(aggregate(by_length[,myvar],by=list(by_length$Length),mean)$x),
              " (",round(aggregate(by_length[,myvar],by=list(by_length$Length),sd)$x),")"))
}
rownames(birthorder.tab)<-levels(by_length$Length)
colnames(birthorder.tab)<-c("Word","Phoneme","NLD")
apa_table(birthorder.tab,caption="NWR measured in whole-word scores, phoneme-based scores, and normalized Levenshtein Distance, separately for the four stimuli lengths.")
```
